Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

We introduce five types of effectiveness \cite{BFM-ic17} that will be essential for solving various resilience problems in different models of WSTS.
\noindent
%			{\bf On effectivity.}
We say that a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{itemize}
\item {\em effective} if there exists a pair of algorithms
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em post-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x \in X$.
%
\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
to the encoding of the ideal $\downarrow s$ is computable and inclusion of ideals is decidable.
%
\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\end{itemize}
%Most of usual WSTS enjoy most of all effectivities.

@@

\begin{theorem}\label{indec WSTS with dcs}
Resilience for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
\end{theorem}

\begin{proof}
\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?...... preuve pas claire, à réécrire}
This stems from the fact that it is undecidable for Minsky machines with more than one counter,
which can be simulated by WSTS.
Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
Remark that in the construction, from any input with at least $1$ on its third counter, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in a $3$-counter Minsky machine
from any input with at least $1$ on its third counter is undecidable. 
Hence {\sc Resilience} is undecidable for Minsky machines.

Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
Reset-VASS extend the basic VASS model with special “reset
transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.

>>>>

We introduce five types of effectiveness \cite{BFM-ic17} that will be essential for solving various resilience problems in different models of WSTS.
\noindent
%			{\bf On effectivity.}
We say that a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{itemize}
\item {\em effective} if there exists a pair of algorithms
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em post-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x \in X$.
%
\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
to the encoding of the ideal $\downarrow s$ is computable and inclusion of ideals is decidable.
%
\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\end{itemize}
Most of usual WSTS are post-effective and ideally-effective.

@@

\begin{theorem}\label{indec WSTS with dcs}
{\sc Resilience} for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
\end{theorem}
\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?}

\begin{proof}
\alain{preuve pas claire, à réécrire}
This stems from the fact that it is undecidable for Minsky machines with more than one counter,
which can be simulated by WSTS.
Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
Remark that in the construction, from any input with at least $1$ on its third counter, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in a $3$-counter Minsky machine
from any input with at least $1$ on its third counter is undecidable. 
Hence {\sc Resilience} is undecidable for Minsky machines.

Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}.

%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
